The SMT Competition 2015--2018

### 2019-04-14 2018/report/SMT-COMP.tex
Report: Restructre section 4 to add more detailed scoring information.

The SMT Competition consists of several tracks.  The \emph{\maintrack}, which
has been a staple of the competition since its
inception in 2005, features non-incremental benchmarks.  The
\emph{\apptrack}, which has been part of the competition since
2011~\cite{BDdMOS13}, features incremental benchmarks.  These
benchmarks are passed to each solver one command at a time by a trace
executor, which waits for the solver's response before passing the
next benchmark command (to prevent look-ahead behavior).  The
\emph{\ucoretrack} uses \maintrack benchmarks that are known to
be unsatisfiable, and solvers are evaluated on their ability to find
small unsatisfiable cores (i.e., a small but still unsatisfiable
subset of the benchmark's assertions).  The \ucoretrack had first
been part of the competition in 2012~\cite{CGBD12}, but was then
discontinued partly because of the competition's move to StarExec and
lacking tool support.  It was re-introduced as an experimental track
in 2016, and has been a regular competition track again since~2017.

Each track is further split into (logic) \emph{divisions}, where each
division corresponds to a logic from the SMT Library.
Table~\ref{table:divisions} shows the number of divisions in each
track since~2015.  Participants can choose which tracks and divisions
they want to enter, and competition results are reported separately
for each track and division.  Thus, competing solvers do not need support
all logics, and may or may not support features such as incrementality
or unsat-core generation.

There are slightly fewer divisions than logics in the SMT Library
(Table~\ref{table:smtlib}) because not every logic contains benchmarks
eligible for the competition.  Some benchmarks typically have to be
excluded for syntactic reasons: in 2015 and 2016, 11,966 benchmarks
were excluded because they contained partial or underspecified
operations (e.g., bit-vector division) whose precise semantics were
being debated at the time.  Additionally in~2016, one benchmark was
excluded because it contained incorrect status information.  In~2017,
1,104 benchmarks were excluded because their logic had been classified
incorrectly, and 8~benchmarks were excluded because of incorrect or
missing status information.  Moreover, divisions containing datatypes
({DT})---which were new in~2017---were excluded from the Unsat-Core
track because of lacking tool support.  In~2018, 3,603~benchmarks were
excluded because they contained named terms (a feature of the SMT-LIB
language that is not permitted in competition benchmarks), and
5~benchmarks were excluded because of missing status information.

@@

SMT Competitions until~2014 employed a pseudo-random selection process
to further cull the number of benchmarks used in the competition.
However, a 2013 evaluation of SMT-COMP and SMT-LIB found that this
``significantly lessens the ability of a competition to determine
`best' solvers.''~\cite{CSW15} Together with the increased
computational resources available to the competition since its move to
StarExec, this finding prompted a paradigm shift.  Since 2015, the
competition is evaluating all solvers on all eligible benchmarks.
This change had an unintended side effect: the pseudo-random selection
process had been purposely constrained to choose benchmarks from
different benchmark families (that is, groups of benchmarks with
similar characteristics).  Since there are vast differences in size
between benchmark families in the SMT Library, the competition
since~2016 uses a scoring method that assigns a weight to each
\maintrack benchmark in order to de-emphasize large benchmark
families~\cite{rules18}.  A benchmark's weight is computed as $(1 +
\ln F)/F$, where~$F\geq 1$ is the size of the benchmark's family.  The
log scaling seems a reasonable compromise between linearly combining
numbers of benchmarks, which would overemphasize large families, and
simply averaging the fraction of benchmarks solved, which would
overemphasize small families.

Another paradigm shift concerns the treatment of benchmarks with
unknown status.  Because of the difficulty to judge solver responses
for such benchmarks, SMT Competitions until 2015 only used benchmarks
whose status was known (i.e., that had been solved previously by
existing SMT solvers).  In 2015, 30,717 non-incremental benchmarks
from the SMT Library were not used for the competition because their
status was unknown.  However, this approach was criticized for
imitating existing solvers rather than truly innovating.  In~2016,
the SMT Library contained 29,724 non-incremental benchmarks with
unknown status; solver performance on these was evaluated but reported
separately.  Since 2017, all non-incremental benchmarks are eligible
for the competition's \maintrack regardless of status, and they
directly affect the competition results.  Solver responses for such
benchmarks are assumed to be correct, unless there is a disagreement
between two or more (otherwise sound e.g. on problems with known status) solvers.  Benchmarks with
disagreements are removed from the competition results.  This conservative approach was chosen over a voting system as this situation is expected to be rare as the SMT-LIB organisers go to a significant effort prior to each competition to update the statuses of problems. A manual investigation is therefore preferred over an automated resolution. Disagreements are reported to the solvers involved and in all cases have led to bugs in solvers being identified and fixed. 
In 2016,
there were disagreements on 79 benchmarks; in 2017, only one benchmark
was removed; and in 2018, there were no disagreements between
otherwise sound solvers.

@@

>>>>

The SMT Competition consists of several tracks.  Each track is further split
into (logic) \emph{divisions}, where each division corresponds to a logic from
the SMT Library.  As indicated in
Tables~\ref{table:smtlib}--\ref{table:divisions}, there are slightly fewer
divisions (Table~\ref{table:divisions}) than logics in the SMT Library
(Table~\ref{table:smtlib}) because not every logic contains benchmarks eligible
for the competition.

\paragraph{Tracks.}
The \emph{\maintrack}, which
has been a staple of the competition since its
inception in 2005, features non-incremental benchmarks.  The
\emph{\apptrack}, which has been part of the competition since
2011~\cite{BDdMOS13}, features incremental benchmarks.  These
benchmarks are passed to each solver one command at a time by a trace
executor, which waits for the solver's response before passing the
next benchmark command (to prevent look-ahead behavior).  The
\emph{\ucoretrack} uses \maintrack benchmarks that are known to
be unsatisfiable, and solvers are evaluated on their ability to find
small unsatisfiable cores (i.e., a small but still unsatisfiable
subset of the benchmark's assertions).  The \ucoretrack had first
been part of the competition in 2012~\cite{CGBD12}, but was then
discontinued partly because of the competition's move to StarExec and
lacking tool support.  It was re-introduced as an experimental track
in 2016, and has been a regular competition track again since~2017.

\paragraph{Divisions.}
Table~\ref{table:divisions} shows the number of divisions in each
track since~2015.  Participants can choose which tracks and divisions
they want to enter, and competition results are reported separately
for each track and division.  Thus, competing solvers do not need to support
all logics, and may or may not support features such as incremental solving
or unsat-core generation.

\paragraph{Eligible Benchmarks.}
Some benchmarks typically have to be
excluded for syntactic reasons: in 2015 and 2016, 11,966 benchmarks
were excluded because they contained partial or underspecified
operations (e.g., bit-vector division) whose precise semantics were
being debated at the time.  Additionally in~2016, one benchmark was
excluded because it contained incorrect status information.  In~2017,
1,104 benchmarks were excluded because their logic had been classified
incorrectly, and 8~benchmarks were excluded because of incorrect or
missing status information.  Moreover, divisions containing datatypes
({DT})---which were new in~2017---were excluded from the Unsat-Core
track because of lacking tool support.  In~2018, 3,603~benchmarks were
excluded because they contained named terms (a feature of the SMT-LIB
language that is not permitted in competition benchmarks), and
5~benchmarks were excluded because of missing status information.

@@

\paragraph{Benchmark Selection.}
SMT Competitions until~2014 employed a pseudo-random selection process
to further cull the number of benchmarks used in the competition.
However, a 2013 evaluation of SMT-COMP and SMT-LIB found that this
``significantly lessens the ability of a competition to determine
`best' solvers''~\cite{CSW15}. Together with the increased
computational resources available to the competition since its move to
StarExec, this finding prompted a paradigm shift.  Since 2015, the
competition is evaluating all solvers on all eligible benchmarks.
This change had an unintended side effect: the pseudo-random selection
process had been purposely constrained to choose benchmarks from
different benchmark families (that is, groups of benchmarks with
similar characteristics).

\paragraph{Benchmarks with Unknown Status.}
Another paradigm shift concerns the treatment of benchmarks with
unknown status.  Because of the difficulty to judge solver responses
for such benchmarks, SMT Competitions until 2015 only used benchmarks
whose status was known (i.e., that had been solved previously by
existing SMT solvers).  In 2015, 30,717 non-incremental benchmarks
from the SMT Library were not used for the competition because their
status was unknown.  However, this approach was criticized for
imitating existing solvers rather than truly innovating.  In~2016,
the SMT Library contained 29,724 non-incremental benchmarks with
unknown status; solver performance on these was evaluated but reported
separately.  Since 2017, all non-incremental benchmarks are eligible
for the competition's \maintrack regardless of status, and they
directly affect the competition results.  Solver responses for such
benchmarks are assumed to be correct, unless there is a disagreement
between two or more (otherwise sound e.g. on problems with known status) solvers.  Benchmarks with
disagreements are removed from the competition results.  This conservative approach was chosen over a voting system as this situation is expected to be rare as the SMT-LIB organisers go to a significant effort prior to each competition to update the statuses of problems. A manual investigation is therefore preferred over an automated resolution. Disagreements are reported to the solvers involved and in all cases have led to bugs in solvers being identified and fixed. 
In 2016,
there were disagreements on 79 benchmarks; in 2017, only one benchmark
was removed; and in 2018, there were no disagreements between
otherwise sound solvers.

@@

\subsection{Scoring}

Since there are vast differences in size
between benchmark families in the SMT Library, the competition
since~2016 uses a scoring method that assigns a weight to each
\maintrack benchmark in order to de-emphasize large benchmark
families~\cite{rules18}.  A benchmark's weight is computed as $(1 +
\ln F)/F$, where~$F\geq 1$ is the size of the benchmark's family.  The
log scaling seems a reasonable compromise between linearly combining
numbers of benchmarks, which would overemphasize large families, and
simply averaging the fraction of benchmarks solved, which would
overemphasize small families.
